---
title: "Getting Started with Regression in R"
output:
  pdf_document: default
  html_notebook: default
editor_options:
  chunk_output_type: console
---

How much variable x interfer with variable y is possible to predict using regression. 
Using linear and no-linear regression using polynomias and spines. 


# load library
```{r}
library("readxl")

library(splines) # for complex datasets


```

# load dataset
```{r}
# dummy dataset to analyse
data_dummy <- read_xlsx("C:/Regression with R/01_example.xlsx")

```

# graph analysis
```{r}

plot(data_dummy$x,data_dummy$y, col="blue", main="Example  Graph")
grid(nx= 12, ny=12, col = "lightgray", lty = "dotted", lwd = par("lwd"), equilogs = TRUE)

```
The result seems to follow a non-linear pattern

# model guessing that is second degree polynomial
```{r}

my_model <- lm(data_dummy$y ~ poly(data_dummy$x, 2))

# but you can test another alternatives
my_model_linear <- lm(data_dummy$y ~ poly(data_dummy$x, 1))

my_model_degree_20 <- lm(data_dummy$y ~ poly(data_dummy$x, 20))

my_model_spline <- lm(data_dummy$y ~ bs(data_dummy$x)) # Here, bs is the base function. Use the parameters knots and df to make the function smoother or curvier.



```

# check polynommial result
```{r}

plot(data_dummy$x,data_dummy$y, col="blue", main="Example  Graph")
grid(nx= 12, ny=12, col = "lightgray", lty = "dotted", lwd = par("lwd"), equilogs = TRUE)
lines(data_dummy$x, predict(lm(data_dummy$y ~ poly(data_dummy$x, 2))))

```


# check alternative linear
```{r}

plot(data_dummy$x,data_dummy$y, col="blue", main="Example  Graph")
grid(nx= 12, ny=12, col = "lightgray", lty = "dotted", lwd = par("lwd"), equilogs = TRUE)
lines(data_dummy$x, predict(my_model_linear <- lm(data_dummy$y ~ poly(data_dummy$x, 1))))

```

# check alternative degree 20
```{r}

plot(data_dummy$x,data_dummy$y, col="blue", main="Example  Graph")
grid(nx= 12, ny=12, col = "lightgray", lty = "dotted", lwd = par("lwd"), equilogs = TRUE)
lines(data_dummy$x, predict(my_model_degree_20 <- lm(data_dummy$y ~ poly(data_dummy$x, 20))))

```

# check model spline
```{r}

plot(data_dummy$x,data_dummy$y, col="blue", main="Example  Graph")
grid(nx= 12, ny=12, col = "lightgray", lty = "dotted", lwd = par("lwd"), equilogs = TRUE)
lines(data_dummy$x, predict(my_model_spline <- lm(data_dummy$y ~ bs(data_dummy$x))))


```


# plot models
```{r}

plot(my_model)

plot(my_model_linear)

plot(my_model_degree_20)

plot(my_model_spline)

```


# t test --This test is going to compare their means, assuming they both are under a normal distribution.
```{r}

t.test(data_dummy$y, predict(my_model))

t.test(data_dummy$y, predict(my_model_linear))

t.test(data_dummy$y, predict(my_model_degree_20))

t.test(data_dummy$y, predict(my_model_spline))


```


Fonte: 
https://www.datasciencecentral.com/profiles/blogs/getting-started-with-regression-in-r
